### 2020-07-13 response/response-demography-3.tex

[Response:]

\item There is actually no data on sex selection and in a sense the data
is backed out using evidence on birth spacing, institutional information
on the availability of technology that enables sex selection (ultra
sound techniques) and government regulations. To do this the author
subdivides the time into four segments: 1972--1984, 1985--1994, 1995--2004
and 2005--2016. I understand the logic of the first two breaks 1984 and
1994, but what is the logic of the third break in 2004? This needs to be
made clearer.

[Response:]

The 2004 break is an attempt to understand whether we are observing the beginning of
a reversal in son preference and the use of sex selection as hypothesized by part of
the prior literature.
In the ``Data'' section, I have split the paragraph on the periods into two and expanded 
the discussion of the motivation for the breaks, including adding two additional
references arguing that we might see the reversal.
Clearly, the 2004/05 break is somewhat arbitrary, but does have the advantage of making 
the periods close to the same length (there are relatively few birth intervals that begin 
in 2015 and 2016).

### 2020-07-13 paper/sex_selection_spacing-ver4.tex
Explain why 2004/05 break in periods (#21)

[TK where to put son preference discussion?]

@@

Direct information on the use of sex selection is not available, so I compare different 
periods, based on the changes in access and legality of prenatal sex 
determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--83, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2016}.
}
Although the use of sex selection increased even after 1994, we may have passed a 
turning point in its use \citep{Das_Gupta2009,Diamond-Smith2015}.
I, therefore, use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection, can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.

@@

In summary, although birth spacing has increased over time with improving economic 
conditions and falling fertility, most of the increases in birth spacing come from the
increased use of sex selection. 
The most substantial increase over the four decades is a full year for the best-educated 
women with two girls, which is the group with the most biased sex ratio.

>>>>

% [TK where to put son preference discussion?]
% %
% \footnote{
% It is also possible that the effect is driven by a stronger son preference, although 
% higher education women have shown declining son preference \citep{bhat03,pande07}.
% }

@@

Direct information on the use of sex selection is not available, so I compare different 
periods, based on the changes in access and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--83, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2016}.
}
Finally, although the use of sex selection increased even after 1994, we may have passed a 
turning point in its use in the mid-2000s with son preference declining
\citep{Das_Gupta2009,Kumar2012,Bongaarts2013,Diamond-Smith2015}.%
\footnote{
Although see also \citep{Navaneetham2011}.
}

I use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The first period covers before sex selection became available and the second the period 
from when sex selection became available until the Prenatal Diagnostic Techniques Act.
I have split the period from 1995 until 2016 into two to understand if there is any 
support for the prior literature's hypothesized reversal in child sex ratios and son 
preference in India.
The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection, can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.

@@

In summary, although birth spacing has increased over time with improving economic 
conditions and falling fertility, most of the increases in birth spacing come from the
increased use of sex selection. 
The most substantial increase over the four decades is a full year for the best-educated 
women with two girls, which is the group with the most biased sex ratio.
There is no clear evidence for or against a reversal in the use of sex selection, with 
some showing increases in sex ratios between the last two periods, others little change, 
and some a decline.
